A weighted sampling algorithm for the design of RNA sequences with targeted secondary structure and nucleotides distribution

### 2013-03-12 ISMB/conclusion_ISMB.tex
Typos + Changed sampling figure

In this article, we described a novel algorithm, \ourprog, for the RNA design problem, i.e. the design of an RNA sequence adopting a predefined secondary structure as minimal free-energy fold.
Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor (e.g. $\sim$28 mins for 100 candidate sequences for a $\sim$1500nts 16s rRNA). We evaluated the method on a benchmark composed of target secondary structures extracted from the \RNASTRAND database. We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good to excellent entropy within designed candidates.
Finally, we implemented an hybrid approach, using the \RNAinverse software as a post-processing step for unpaired regions. This approach greatly increased the success rate of the method, allowing for the design of highly diverse candidates for almost all of the structures in our benchmark, while largely preserving the targeted \GCContent.

>>>>

In this article, we described a novel algorithm, \ourprog, for the RNA secondary structure design problem, i.e. the design of an RNA sequence adopting a predefined secondary structure as its minimal free-energy fold. Implementing a global sampling approach, it optimizes affinity towards the target secondary structure, while granting the user full control over the \GCContent of the resulting sequences.
This extended control does not necessarily induce additional computational demands, and we showed the linear complexity of both the preprocessing stage and the generation of candidate sequences for the design, allowing for the design of larger and more complex secondary structures in a matter of minutes on a single processor (e.g. $\sim$28 mins for 100 candidate sequences for a $\sim$1500nts 16s rRNA). We evaluated the method on a benchmark composed of target secondary structures extracted from the \RNASTRAND database. We observed good overall success rate, with the notable exception of very low targeted \GCContent ($10\%$), and a good to excellent entropy within designed candidates.
Finally, we implemented an hybrid approach, using the \RNAinverse software as a post-processing step for unpaired regions. This approach greatly increased the success rate of the method, allowing for the design of highly diverse candidates for almost all of the structures in our benchmark, while largely preserving the targeted \GCContent.

### 2013-03-12 ISMB/results_ISMB.tex

\highlight{
\subsection{Global sampling vs Local search vs Glocal approach}
To conclude this study, we estimate the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD). To our knowledge, \RNASSD, beside \ourprog, is the only software that implements an explicit control of the \GCContent.

@@

Next, we show in Figure~\ref{fig:benchmark_RNASSD} the average sequence identity achieved by the various methods. Our results show that at extremely low {\GCContent}s (i.e. 10\%), \ourprog slightly outperforms \RNASSD while this advantage becomes less evident when the \GCContent increases. Our experiments on higher {\GCContent}s (i.e. 50\% and above) showed that our glocal strategy and the local search approach perform similarly. Similarly, we did not find any clear evidence that a global, local or glocal approach outperforms others when we compare at the thermodynamical stability of the target structure (data not shown).
}

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
  \label{fig:benchmark_RNASSD}
\end{figure*}

>>>>

%Here, we are interested in estimating the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and local search methodologies (\RNAinverse, \INFORNA, \NUPACK and \frankenstein). 
%=======
\highlight{
\subsection{Global sampling vs Local search vs Glocal approach}

To conclude this study, we estimate the impact of the design methodology on the performances. More precisely, we aim to determine the merits of a global sampling approach (\ourprog), compared to a glocal procedure (\ourprog + \RNAinverse) and a local search methodology (\RNASSD). To our knowledge, \RNASSD, beside \ourprog, is the only software that implements an explicit control of the \GCContent.

@@

Next, we show in Figure~\ref{fig:benchmark_RNASSD} the average sequence identity achieved by the various methods. Our results show that at extremely low {\GCContent}s (i.e. 10\%), \ourprog slightly outperforms \RNASSD while this advantage becomes less evident when the \GCContent increases. Our experiments on higher {\GCContent}s (i.e. 50\% and above) showed that our glocal strategy and the local search approach perform similarly. Similarly, we did not find any clear evidence that a global, local or glocal approach outperforms others when we compare at the thermodynamical stability of the target structure (data not shown).
}
%>>>>>>> af029e5428ff6e8980ef29cd0002a6fd72501bd9

%\begin{figure*}[ht!]
%  \centering
%  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
%  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
%\end{figure*}
\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{Figures/seq_identity_10_30}
  \caption{Sequence identity of \ourprog and \RNAinverse for $10$ and $30\%$ of \GC.}
  \label{fig:benchmark_RNASSD}
\end{figure*}
